AGB Stars in the IR: Physical Characteristics, Galactic Spatial Distribution, and Chemical Impact of the Milky Way AGB Population

### 2015-04-01 paper1/paper1_criteria.tex
updating the paper

\section{Object Selection Criteria}
\label{sec:criteria}
We made color-color criteria to create a catalog of AGB candidates. We kept it simple to maximize the number of objects we'd retain while minimizing contamination from non-AGB objects.

>>>>

\section{Object Selection Criteria}
\label{sec:criteria}
We create color-color criteria to generate a catalog of AGB candidates. We seek to maximize AGB completeness while minimizing contamination from non-AGB objects to beneath the 1\% level.

The color-color criteria for AGB selection are as follows:
\begin{align} 
(J-K_s) > 1.1\label{eq:criteria1}\\
(W2-W3) > 0.3\label{eq:criteria2}\\
(W3-W4) < -0.83(W2-W3) + 3.37\label{eq:criteria3}
\end{align}
The criteria in (\ref{eq:criteria1}) and (\ref{eq:criteria2}) are primarily concerned with rejecting objects from the stellar locus, and other stars whose NIR spectra are dominated by the Rayleigh-Jeans tail (many covered by the sample of galaxies). (\ref{eq:criteria3}) also rejects stars from the stellar locus, but primarily functions to remove IR-bright extragalactic sources.

\noindent Sample completeness $\eta$ is defined as
\begin{eqnarray*}
\eta &=& \frac{N - n_\text{missed}}{N}
\end{eqnarray*}
where $N$ is the total number of objects in the sample, and $n_\text{missed}$ is the number of objects outside of the applied boundaries. Figure~\ref{fig:completeness} shows the distribution of sample completeness amongst AGB sources after the application of the above criteria. The vast majority (79.07\%) of Galactic AGB stars is recovered after criteria (\ref{eq:criteria1}), (\ref{eq:criteria2}), and (\ref{eq:criteria3}) are applied. The largest losses occur at the edges of the Galactic disk ($\lvert b\rvert\approx10^\circ$) and in regions of high stellar number density both in the Galaxy and the Magellanic Clouds. In the color-color diagrams, selection completeness degrades near the borders of the selction area, as objects that straddle these boundaries may exhibit enough emission in other color-color spaces to be removed by our criteria. Of the remaining sample of 5,709 objects, 52.94\% lie within the disk ($\lvert b\rvert<10^\circ$).

\begin{figure}[h]
\includegraphics[width=6.7in]{figs/completeness_map.pdf}
\caption{Sample selection completeness maps in ($l,b$) space (\emph{top}) and the color-color space of our selection criteria (\emph{bottom}). Color scale shows the completeness fraction per bin, with 4-deg$^2$ bins on the Galactic map and 0.1 dex bins on each axis of the color-color diagrams. Selection criteria are shown as dashed lines.\label{fig:completeness}}
\end{figure}

\cite{2013sdmm.book.....I} defines sample contamination as
\begin{eqnarray*}
\epsilon &=& \frac{n_\text{spurious}}{n_\text{selected}}.
\end{eqnarray*}
where $n_\text{spurious}$ is the number of spurious sources and $n_\text{selected} = N + n_\text{spurious}$. The contamination map is shown in Figure~\ref{fig:contamination}. Most bins in Figure~\ref{fig:contamination} exhibit 0\% contamination, with the overall contamination level at 0.38\%. What contaminants do remain exist primarily at the very fringes of the AGB star distribution in ($J-K_S$) vs ($W2 - W3$) space and near the redder boundary in ($W3-W4$) vs ($W2 - W3$), where bluer extragalactic sources creep into the selection region. The results of the applied critiera on both the collective AGB and contaminant samples are summarized in Table~\ref{tab:completeness}.

\begin{figure}[h]
\includegraphics[width=6.7in]{figs/contamination_map.pdf}
\caption{Sample contamination in Galactic ($l,b$) space (\emph{top}) and color-color space (\emph{bottom}). Boundaries and binning the same as Fig~\ref{fig:completeness}.\label{fig:contamination}}
\end{figure}

%One large issue with the boundaries are the O-rich AGB stars.  Either the objects from the OGLE O-rich AGB star catalog were mostly mis-matched to WISE, taking on colors of the stellar locus, or the issue is more physical.  It could be that the warm O-rich AGB photospheres are still visible and not heavily enshrouded in dust, thus appearing similar to Main Sequence stars in the NIR. Previous work on warm O-rich AGB stars has shown that their circumstellar shells are not prominent, and their NIR photometry reflects the Rayleigh Jeans tail of a cool 2000-4000K blackbody \citep{1974ApJ...189...89D}.

\begin{table}[h]
    \begin{center}
        \caption{Sample Selection Completeness and Contamination}
        \scalebox{0.85}{
            \begin{tabular}{l c c c c c c}
		\hline
		Population & SIMBAD AGB* & C* & Mira & OH/IR & S* \\ 
		Completeness & 89.62\% & 72.11\% & 95.62\% & 39.53\% & 22.31\% \\ 
		\hline
		Population & MACHO seq1 & seq2 & seq3 & seq4 \\ 
		Completeness & 88.45\% & 81.08\% & 28.77\% & 14.75\% \\ 
		\hline
		Population & OGLE-III C-rich & O-rich & \textbf{All AGB Stars} \\ 
		Completeness & 73.09\% & 70.68\% & \textbf{79.07\%} \\ 
		\hline
		Population & DR12 SSPP & DR7 LRG & Galaxies & QSO & AGN \\ 
		Contamination & 0.56\% & 0.00\% & 0.00\% & 0.07\% & 	0.00\% \\ 
		\hline

                \label{tab:completeness}
            \end{tabular}}    
    \end{center}
\end{table}

### 2015-04-01 paper1/paper1_data.tex
updating the paper

\section{Data Sources and Reduction}
\label{sec:data}
\subsection{Data Sources}
In order to analyze the physical distribution of \agb\, stars in the IR, we require reliable identifications of known \agb, stars to generate color-color selection criteria. We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}. \ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II. O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in \cite{2009AcA....59..239S,2011AcA....61..217S}. Data reduction techniques are described in \cite{2008AcA....58...69U}. The resulting samples yielded 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich). From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars). The sample of AGB stars from \simbad\, was obtained by querying for all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars. The total sample of \agb\, stars is 100,393.

@@

In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}. The initial \wise All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m. Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$]. The positions of objects in the \wise\, catalog were calibrated to the \twomass\, point source catalog. The 3-band cryogenic data release contains data from $W1$, 2, and 3, and surveyed $30\%$ of the sky between August and October 2010. During the 3-band cryogenic survey, $W1$ and $W2$ operated with nearly the same sensitivity as during the full survey. Warming of the telescope reduced sensitivty in $W3$ and fully saturated $W4$. The NEOWISE post-cryogenic data release contains $W1$ and $W2$ measurements, with sensitivities close to those obtained during the full cryogenic phase. During this phase, \wise\, surveyed $70\%$ of the sky. Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR. The overall total number of sources compiled into \allwise\, totals over 747.6 million.

@@

\begin{figure}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
\caption{\emph{Left:} Known AGB stars after data reduction. \emph{Right:} All contaminants after data reduction.\label{fig:distros}}
\end{figure}

>>>>

\section{Data Sources and Reduction}
\label{sec:data}
\subsection{Data Sources}
In order to analyze the physical distribution of \agb\, stars in the IR, we require reliable identifications of known \agb, stars to generate color-color selection criteria. We select \agb\, stars from three source catalogs: the {\it Optical Gravitational Lens Experiment-III Variable Star Catalog} \citep[\ogle,][]{2008AcA....58...69U,2009AcA....59..239S,2011AcA....61..217S}, the {\it MAssive Compact Halo Objects} project \citep[\macho,][]{1997ApJ...482...89A}, and the \simbad\, Astronomical Database \citep{2000A&AS..143....9W}.

\ogle\, photometry for Long-Period Variables (LPVs) in the Small and Large Magellanic Clouds (SMC and LMC respsectively) was obtained between July 2001 and May 2009, with stars in the central 4.5-deg$^2$ of the LMC and SMC having an additional 5 observing seasons of photometry from OGLE-II. O-rich and C-rich \agb\, stars in \ogle\, were photometrically selected using reddening-free Wesenheit magnitudes, described in \cite{2009AcA....59..239S,2011AcA....61..217S}. Data reduction techniques are described in \cite{2008AcA....58...69U}. The resulting samples yielded 46,467 \agb\, stars from the LMC (37,203 O-rich; 9,264 C-rich) and 6,509 stars from the SMC (3,727 O-rich; 2,782 C-rich).

From \macho\, we obtain the sample of SMC, LMC, and Galactic Bulge AGB stars used in \cite{2008AJ....136.1242F} (14,861 stars). Following \cite{2008AJ....136.1242F}, the objects are divided into sequences (seq) 1-4. Sequence 1 primarily contains Miras pulsating in their fundamental modes, whereas Sequences 2-4 contain semi-regular variables in various pulsation modes.

The sample of AGB stars from \simbad\, was obtained by querying all objects classified as C-stars (18,656), S-stars (1,108), OH/IR stars (825), AGB stars (2,359), and Mira variables (9,608), for a total of 32,556 stars. The total sample of \agb\, stars is 100,393. Objects are classified spectroscopically, though by a variety of methods owing to the heterogeneous data housed within \simbad.

@@

In this study, we rely heavily on data from the \allwise\, extension of the \wise\, survey, combining data from the initial All-Sky Data Release, the 3-band cryogenic data release, and the NEOWISE post-cryogenic data release \citep{2013wise.rept....1C}. The initial \wise\,All-Sky Data Release observed the sky between January and August 2010, observing the sky 1.2 times with four detectors, operating at 3.4, 4.6, 12, and 22$\mu$m. Hereon we refer to \allwise\, photometric bands at [3.4$\mu$m/4.6$\mu$m/12$\mu$m/22$\mu$m] as [$W1/W2/W3/W4$]. The positions of objects in the \wise\, catalog were calibrated to the \twomass\, point source catalog. The 3-band cryogenic data release contains data from $W1$, 2, and 3, and surveyed $30\%$ of the sky between August and October 2010. During the 3-band cryogenic survey, $W1$ and $W2$ operated with nearly the same sensitivity as during the full survey. Warming of the telescope reduced sensitivty in $W3$ and fully saturated $W4$. The NEOWISE post-cryogenic data release contains $W1$ and $W2$ measurements, with sensitivities close to those obtained during the full cryogenic phase. During this phase, \wise\, surveyed $70\%$ of the sky. Data products from the post-cryogenic release included updated instrumental, astrometric, and photometric calibrations and reduction algorithms, resulting in much lower SNR. The overall total number of sources compiled into \allwise\, totals over 747.6 million.

@@

\begin{figure}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color1.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color2.pdf}
\includegraphics[width=7in]{figs/agbs_contaminants_color_color3.pdf}
\caption{Logarithmic number densities for objects in \wise\, and \twomass\, color-color space, binned in 0.1 dex on each axis. \emph{Left:} The combined AGB sample matched to ALLWISE. \emph{Right:} The combined contaminant sample.\label{fig:distros}}
\end{figure}
